---
title: "Basic GO Analysis with mogo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic GO Analysis with mogo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## osgo

`osgo` is a package for generating GO analysis using _O.sativa_ genes only. You can provide the main function with the list of genes of interest and it does (most of) the rest!


## Preparation

The first step is to load our gene expression data file and get a simple list of genes for GO analysis

### Data loading

First read in your gene expression file. You can do that with `read_csv()`. Minimally it should contain the gene ID, the log fold change and the $p$-value.


```{r setup, message=FALSE}
library(readr)
library(dplyr)
library(here)

gene_expression <- read_csv(here("inst", "extdata","sample_gene_expression.csv"))
gene_expression
```

### Filter as required

We can now filter the genes to select only the ones with e.g $p <= 0.05$, using `filter()`

```{r, message=FALSE}
filtered_gene_expression <- filter(gene_expression, p.adj <= 0.05)
filtered_gene_expression
```

### Extract the gene id column

We can now extract the `gene_id` column using the `$` syntax

```{r, message=FALSE}
gene_ids <- filtered_gene_expression$gene_id
gene_ids
```

## GO enrichment

The GO enrichment is done in the `osgo` package. Load that and use the `do_enrich()` function, passing it the vector of `gene_ids` to calculate the enrichment. **Note that you must here specify which _O.sativa_ group you use.** Use the `which` argument with either `"indica"` or `"japonica"`. 

```{r, message=FALSE, warning=FALSE}
library(osgo)
enrich <- do_enrich(gene_ids, which = "japonica")
enrich
```


As you can see the `enrich` object has a lot of information in it. The result table can be extracted using `as.data.frame()` to convert the `enrich` to a data.frame (note `glimpse` is a helpful function for printing out big dataframes).


```{r}
result_table <- as.data.frame(enrich)
glimpse(result_table)
```

You can save the result to an excel-compatible csv file with, `write_csv()`

```{r, eval=FALSE}
write_csv(result_table, "my_GO_results.csv")
```

## Plotting

The `enrich` object is from the package `ClusterProfiler` and can be used directly in most of it's plot types. See them at the `ClusterProfiler` page [http://yulab-smu.top/clusterProfiler-book/chapter12.html#bar-plot](http://yulab-smu.top/clusterProfiler-book/chapter12.html#bar-plot)

### Barplot

Works ok, but the text can make it a bit unwieldy

```{r,fig.width=24, fig.height=8}
library(clusterProfiler)
library(enrichplot)
barplot(enrich, showCategory=5)
```

### Dotplot

Similar

```{r, fig.width=18, fig.height=8}
dotplot(enrich, showCategory=5)
```

### UpSetPlot

These are like a really sophisticated Venn/Euler diagram. [https://jku-vds-lab.at/tools/upset/](https://jku-vds-lab.at/tools/upset/)

```{r,fig.width=18, fig.height=8}
upsetplot(enrich)
```

### GoPlot

Plots from the package `GoPlot` such as bubble plots can be made, but need you to convert the `enrich` object to a DAVID compatible object, and make an accessory data.frame of the expression information. Use `enricher_to_david()` for the first part, and the relevant columns of the gene expression data you created at the beginning for the second part. **Note that for david conversion you must here specify which _O.sativa_ group you use.** Use the `which` argument with either `"indica"` or `"japonica"`.

```{r}
david <- enricher_to_david(enrich, which = "japonica")
expr_info <- data.frame(
  ID = filtered_gene_expression$gene_id,
  logFC = filtered_gene_expression$log2fc
)
```

Then you can convert that to the data format needed for `GoPlot`

```{r}
library(GOplot)
circ <- circle_dat(david, expr_info)
```

### Go Bar

A different sort of GO barchart

```{r, fig.height=12, fig.width=12}
GOBar(subset(circ, category == 'BP'))
```

### Go Bubble

A plot with bubbles

```{r, eval=FALSE, ig.height=12, fig.width=12}
GOBubble(circ, labels=3)
```



